---
title: "Fintech"
author: "Derik Suria"
output: html_document
date: "2022-09-14"
---

## Loading Packages

```{r setup, include=FALSE}
install.packages("tidyverse")
install.packages("padr")
install.packages("zoo")
install.packages("lubridate")
install.packages("modelr")
```

```{r}
library(tidyverse)
library(padr)
library(zoo)
library(lubridate)
library(modelr)
```

## Reading csv Files

```{r reading files}
fred5ybe <- read_csv("5BE.csv") %>%
  mutate(Date = parse_date(Date, format = "%m/%d/%y"))
fred5ybe
```

```{r}
fred5yfl <- read_csv("5FL.csv") %>%
  mutate(Date = parse_date(Date, format = "%m/%d/%y"))
fred5yfl
```

```{r}
fred10ybe <- read_csv("10BE.csv") %>%
  mutate(Date = parse_date(Date, format = "%m/%d/%y"))
fred10ybe
```

```{r}
crypt <- read_csv("crypt.csv") %>%
  rename(index = "S&P Cryptocurrency Broad Digital Market Index (USD)",
         date = "Effective date") %>%
  select(date, index) %>%
  mutate(date = parse_date(date, format = "%m/%d/%y"))
```

```{r data from new gold}
gold <- read_csv("gold.csv") %>%
  mutate(Date = parse_date(Date, format = "%m/%d/%y")) %>%
  select(Date, GOLD)
gold
```

```{r ETF}
etf <- read_csv("etf.csv") %>%
  select(Date, etf) %>%
  mutate(Date = parse_date(Date, format = "%m/%d/%y"))
etf
```

## Filling Missing Dates using pad()

```{r}
fiveBE <- pad(fred5ybe) #https://statisticsglobe.com/insert-rows-for-missing-dates-in-r
fiveFL <- pad(fred5yfl)
tenBE <- pad(fred10ybe)
cry <- pad(crypt)
goldc <- pad(gold)
etfc <- pad(etf)
```

## Interpolating Missing Values using na.approx()

```{r}
d5 <- data_new %>%
  rename(ifl = "5-Year Breakeven Inflation Rate, Percent, Daily, Not Seasonally Adjusted") %>%
  mutate(ifl = na.approx(ifl)) #https://www.statology.org/r-interpolate-missing-values/, https://stats.stackexchange.com/questions/107505/interpolating-time-series
d5 <- data2 %>% 
  select(T5YIE, ifl)
d5
```

```{r}
fiveBEclean <- fiveBE %>%
  rename(inf = "5BE") %>%
  mutate(inf = na.approx(inf)) 
```

```{r}
fiveFLclean <- fiveFL %>%
  rename(inf = "5FL") %>%
  mutate(inf = na.approx(inf)) 
```

```{r}
tenBEclean <- tenBE %>%
  rename(inf = "10BE") %>%
  mutate(inf = na.approx(inf)) 
```

```{r}
cryptclean <- cry %>%
  mutate(index = na.approx(index)) 
```

```{r}
goldclean <- goldc %>%
  mutate(GOLD = na.approx.default(GOLD, rule = 2)) #https://stackoverflow.com/questions/7317607/interpolate-na-values-in-a-data-frame-with-na-approx
```

```{r}
etfclean <- etfc %>%
  mutate(etf = na.approx(etf)) 
```

## Exporting Clean Data

```{r}
write_csv(fiveBEclean, "/Users/oneinchdong/Desktop/ECON181/5yrBEclean.csv")
write_csv(tenBEclean, "/Users/oneinchdong/Desktop/ECON181/10yrBEclean.csv")
write_csv(fiveFLclean, "/Users/oneinchdong/Desktop/ECON181/5yrFLclean.csv")
write_csv(cryptclean, "/Users/oneinchdong/Desktop/ECON181/cryptclean.csv")
write_csv(goldclean, "/Users/oneinchdong/Desktop/ECON181/goldclean.csv")
write_csv(etfclean, "/Users/oneinchdong/Desktop/ECON181/etfclean.csv")
```

# Granger Causality Test

```{r}
install.packages("lmtest")
library (lmtest)
```

```{r}
cause <- read_csv("clean_long.csv")
```

## BTC DR vs Explanatory Variables

**Notation/Notes: grangertest(y \~ x, order = 1)** - This function tests whether time series x is a valuable tool for predicting the values of time series y, conditional on having already accounted for the effects on yt of past values of y (and perhaps of past values of other variables). - Null hypothesis: time series x does not granger-cause time series y - Alternative hypothesis: time series x granger-causes time series y, meaning past values of times series x are useful in predicting values of time series y, implying causation - order = number of lags to utilize where the default value is 1; ie. order = 1 means comparing Sunday vs Monday, order = 2 means comparing Saturday vs Monday, etc. - Interpretation: observe if the p-value is greater than the threshold value of 0.05 (95% CI)-- if greater than fail to reject null, if smaller can reject null - Checking if inflation is a leading indicator

```{r}
grangertest(BTCDR ~ FBE, order = 1, data = cause) # p-value = 0.05791
grangertest(BTCDR ~ FBE, order = 2, data = cause) # p-value = 0.01406
grangertest(BTCDR ~ FBE, order = 3, data = cause) # p-value = 0.03389
grangertest(BTCDR ~ FBE, order = 4, data = cause) # p-value = 0.02118
grangertest(BTCDR ~ FBE, order = 5, data = cause) # p-value = 0.04272
```

```{r}
grangertest(BTCDR ~ FFL, order = 1, data = cause) # p-value = 0.963
grangertest(BTCDR ~ FFL, order = 2, data = cause) # p-value = 0.613
grangertest(BTCDR ~ FFL, order = 3, data = cause) # p-value = 0.02927
grangertest(BTCDR ~ FFL, order = 4, data = cause) # p-value = 0.06439
grangertest(BTCDR ~ FFL, order = 5, data = cause) # p-value = 0.07557
```

```{r}
grangertest(BTCDR ~ TBE, order = 1, data = cause) # p-value = 0.2077
grangertest(BTCDR ~ TBE, order = 2, data = cause) # p-value = 0.4204
grangertest(BTCDR ~ TBE, order = 3, data = cause) # p-value = 0.1726
grangertest(BTCDR ~ TBE, order = 4, data = cause) # p-value = 0.2774
grangertest(BTCDR ~ TBE, order = 5, data = cause) # p-value = 0.3807
```

## ETF DR vs Explanatory Variables

```{r}
grangertest(ETFDR ~ FBE, order = 1, data = cause) # p-value = 1.584e-07
grangertest(ETFDR ~ FBE, order = 2, data = cause) # p-value = 1.11e-07
grangertest(ETFDR ~ FBE, order = 3, data = cause) # p-value = 4.181e-07
grangertest(ETFDR ~ FBE, order = 4, data = cause) # p-value = 3.946e-09
grangertest(ETFDR ~ FBE, order = 5, data = cause) # p-value = 3.091e-11
```

```{r}
grangertest(ETFDR ~ FFL, order = 1, data = cause) # p-value = 0.01946
grangertest(ETFDR ~ FFL, order = 2, data = cause) # p-value = 0.02223
grangertest(ETFDR ~ FFL, order = 3, data = cause) # p-value = 0.02725
grangertest(ETFDR ~ FFL, order = 4, data = cause) # p-value = 0.0001925
grangertest(ETFDR ~ FFL, order = 5, data = cause) # p-value = 0.0002907
```

```{r}
grangertest(ETFDR ~ TBE, order = 1, data = cause) # p-value = 4.03e-05
grangertest(ETFDR ~ TBE, order = 2, data = cause) # p-value = 2.837e-05
grangertest(ETFDR ~ TBE, order = 3, data = cause) # p-value = 6.421e-05 
grangertest(ETFDR ~ TBE, order = 4, data = cause) # p-value = 1.713e-09
grangertest(ETFDR ~ TBE, order = 5, data = cause) # p-value = 9.919e-11
```

## GOLD DR vs Explanatory Variables

```{r}
grangertest(GOLDDR ~ FBE, order = 1, data = cause) # p-value = 0.3948
grangertest(GOLDDR ~ FBE, order = 2, data = cause) # p-value = 0.1834
grangertest(GOLDDR ~ FBE, order = 3, data = cause) # p-value = 0.2919
grangertest(GOLDDR ~ FBE, order = 4, data = cause) # p-value = 0.3419
grangertest(GOLDDR ~ FBE, order = 5, data = cause) # p-value = 0.4375
```

```{r}
grangertest(GOLDDR ~ FFL, order = 1, data = cause) # p-value = 0.5155
grangertest(GOLDDR ~ FFL, order = 2, data = cause) # p-value = 0.4878
grangertest(GOLDDR ~ FFL, order = 3, data = cause) # p-value = 0.51
grangertest(GOLDDR ~ FFL, order = 4, data = cause) # p-value = 0.6032
grangertest(GOLDDR ~ FFL, order = 5, data = cause) # p-value = 0.7127
```

```{r}
grangertest(GOLDDR ~ TBE, order = 1, data = cause) # p-value = 0.2863
grangertest(GOLDDR ~ TBE, order = 2, data = cause) # p-value = 0.4758
grangertest(GOLDDR ~ TBE, order = 3, data = cause) # p-value = 0.6374
grangertest(GOLDDR ~ TBE, order = 4, data = cause) # p-value = 0.5316
grangertest(GOLDDR ~ TBE, order = 5, data = cause) # p-value = 0.6377
```

## CRYPT DR vs Explanatory Variables

```{r}
grangertest(CRYPTDR ~ FBE, order = 1, data = cause) # p-value = 0.1809
grangertest(CRYPTDR ~ FBE, order = 2, data = cause) # p-value = 0.0591
grangertest(CRYPTDR ~ FBE, order = 3, data = cause) # p-value = 0.04573
grangertest(CRYPTDR ~ FBE, order = 4, data = cause) # p-value = 0.0745
grangertest(CRYPTDR ~ FBE, order = 5, data = cause) # p-value = 0.1311
```

```{r}
grangertest(CRYPTDR ~ FFL, order = 1, data = cause) # p-value = 0.4347
grangertest(CRYPTDR ~ FFL, order = 2, data = cause) # p-value = 0.5628
grangertest(CRYPTDR ~ FFL, order = 3, data = cause) # p-value = 0.7783
grangertest(CRYPTDR ~ FFL, order = 4, data = cause) # p-value = 0.4394
grangertest(CRYPTDR ~ FFL, order = 5, data = cause) # p-value = 0.5381
```

```{r}
grangertest(CRYPTDR ~ TBE, order = 1, data = cause) # p-value = 0.2737
grangertest(CRYPTDR ~ TBE, order = 2, data = cause) # p-value = 0.6499
grangertest(CRYPTDR ~ TBE, order = 3, data = cause) # p-value = 0.7761
grangertest(CRYPTDR ~ TBE, order = 4, data = cause) # p-value = 0.5925
grangertest(CRYPTDR ~ TBE, order = 5, data = cause) # p-value = 0.7305
```

# Univariate Linear Regression (without autoregressive specification)

```{r}
btc1 <- lm(BTCDR ~ lagBTC + FBE, data = cause_lagged)
summary(btc1)
btc2 <- lm(BTCDR ~ lagBTC + FFL, data = cause_lagged)
summary(btc2)
btc3 <- lm(BTCDR ~ lagBTC + TBE, data = cause_lagged)
summary(btc3)
```

```{r}
etf1 <- lm(ETFDR ~ lagETF + FBE, data = cause_lagged)
summary(etf1)
etf2 <- lm(ETFDR ~ lagETF + FFL, data = cause_lagged)
summary(etf2)
etf3 <- lm(ETFDR ~ lagETF + TBE, data = cause_lagged)
summary(etf3)
```

```{r}
gold1 <- lm(GOLDDR ~ lagGOLD + FBE, data = cause_lagged)
summary(gold1)
gold2 <- lm(GOLDDR ~ lagGOLD + FFL, data = cause_lagged)
summary(gold2)
gold3 <- lm(GOLDDR ~ lagGOLD + TBE, data = cause_lagged)
summary(gold3)
```

```{r}
crypt1 <- lm(CRYPTDR ~ lagCRYPT + FBE, data = cause_lagged)
summary(crypt1)
crypt2 <- lm(CRYPTDR ~ lagCRYPT + FFL, data = cause_lagged)
summary(crypt2)
crypt3 <- lm(CRYPTDR ~ lagCRYPT + TBE, data = cause_lagged)
summary(crypt3)
```

# Univariate Linear Regression with Autoregressive Specification (t-1)

```{r}
cause_lagged <- cause %>%
  mutate(lagFBE = lag(FBE, n = 1, default = NA),
         lagFFL = lag(FFL, n = 1, default = NA),
         lagTBE = lag(TBE, n = 1, default = NA),
         lagETF = lag(ETFDR, n = 1, default = NA),
         lagGOLD = lag(GOLDDR, n = 1, default = NA))
```

```{r}
btc1as <- lm(BTCDR ~ lagBTC + lagFBE, data = cause_lagged)
summary(btc1as)
btc2as <- lm(BTCDR ~ lagBTC + lagFFL, data = cause_lagged)
summary(btc2as)
btc3as <- lm(BTCDR ~ lagBTC + lagTBE, data = cause_lagged)
summary(btc3as)
```

```{r}
etf1as <- lm(ETFDR ~ lagETF + lagFBE, data = cause_lagged)
summary(etf1as)
etf2as <- lm(ETFDR ~ lagETF + lagFFL, data = cause_lagged)
summary(etf2as)
etf3as <- lm(ETFDR ~ lagETF + lagTBE, data = cause_lagged)
summary(etf3as)
```

```{r}
gold1as <- lm(GOLDDR ~ lagGOLD + lagFBE, data = cause_lagged)
summary(gold1as)
gold2as <- lm(GOLDDR ~ lagGOLD + lagFFL, data = cause_lagged)
summary(gold2as)
gold3as <- lm(GOLDDR ~ lagGOLD + lagTBE, data = cause_lagged)
summary(gold3as)
```

```{r}
crypt1as <- lm(CRYPTDR ~ lagCRYPT + lagFBE, data = cause_lagged)
summary(crypt1as)
crypt2as <- lm(CRYPTDR ~ lagCRYPT + lagFFL, data = cause_lagged)
summary(crypt2as)
crypt3as <- lm(CRYPTDR ~ lagCRYPT + lagTBE, data = cause_lagged)
summary(crypt3as)
```

# Univariate Regression Controlling for the FED Target Rate of 2%

```{r}
cause_controlled <- cause %>%
  mutate(controlFBE = ifelse(FBE > 0.02, 1, 0),
         controlFFL = ifelse(FFL > 0.02, 1, 0),
         controlTBE = ifelse(TBE > 0.02, 1, 0),
         lagcontrolFBE = lag(controlFBE, n = 1, default = NA),
         lagFBE = lag(FBE, n = 1, default = NA),
         lagFFL = lag(FFL, n = 1, default = NA),
         lagTBE = lag(TBE, n = 1, default = NA),
         lagBTC = lag(BTCDR, n = 1, default = NA),
         lagETF = lag(ETFDR, n = 1, default = NA),
         lagCRYPT = lag(CRYPTDR, n = 1, default = NA),
         lagGOLD = lag(GOLDDR, n = 1, default = NA))
```

```{r}
control <- read_csv("control_data.csv")
control %>%
  mutate(DATE = parse_date(DATE, format = "%d/%m/%y"))

control2 <- control %>%
  mutate(controlFBE = ifelse(FBE > 2, 1, 0),
         lagBTC = lag(BTCDR, n = 1, default = NA),
         lagCRYPT = lag(CRYPT, n = 1, default = NA))
```


```{r}
btc1control <- lm(BTCDR ~ lagBTC + FBE + controlFBE, data = control2)
summary(btc1control)
btc2control <- lm(BTCDR ~ lagBTC + FFL + controlFBE, data = control2)
summary(btc2control)
btc3control <- lm(BTCDR ~ lagBTC + TBE + controlFBE, data = control2)
summary(btc3control)
```

```{r}
crypt1control <- lm(CRYPT ~ lagCRYPT + FBE + controlFBE, data = control2)
summary(crypt1control)
crypt2control <- lm(CRYPT ~ lagCRYPT + FFL + controlFBE, data = control2)
summary(crypt2control)
crypt3control <- lm(CRYPT ~ lagCRYPT + TBE + controlFBE, data = control2)
summary(crypt3control)
```

# Monthly Inflation Swaps vs BTC Monthly Returns

```{r}
btc_m <- read_csv("btc_monthly.csv") %>%
  mutate(Date = parse_date(Date, format = "%d/%m/%y"))
btc_m
```

```{r aggregate daily data to monthly}
btc_m$Month <- floor_date(btc_m$Date, "month") #https://www.statology.org/aggregate-daily-data-in-r/
btc_monthly <- btc_m %>%
  group_by(Month) %>%
  summarize(mean = mean(`BTC DR`))
```

```{r}
write_csv(btc_monthly, "/Users/oneinchdong/Desktop/ECON181/btc_monthly_clean.csv")
```





